---
title: "SSU_week_17"
format: html
editor: visual
---

```{r}
set.seed(123)
```

```{r}
#| echo: false
#| results: hide
library(Hmisc)
library(ggplot2)
library(rms)
library(mice)
library(boot)
library(MASS)
library(pim)
library(patchwork)
library(ggpubr)
library(DHARMa)
```

# Check New Analysis Data

```{r}
# new analysis data
dt2 <- read.csv('analytic_data.csv')

# transform data
dt2$daooh30 <- as.numeric(dt2$daooh30)
dt2$daoohr30 <- as.numeric(dt2$daoohr30)
dt2$daooh90 <- as.numeric(dt2$daooh90)
dt2$daoohr90 <- as.numeric(dt2$daoohr90)

dt2$hosp6mo <- as.factor(dt2$hosp6mo)
dt2$troponinilow2 <- as.factor(dt2$troponinilow2)
```

```{r}
colnames(dt2)
```

```{r}
# data summary
table2 <- summaryM(daooh30+daooh90+daoohr30+daoohr90+hosp6mo+sbp+bun+troponinilow2+bnpval ~treatgroup,data = dt2)
html(table2)

sapply(dt2[,c('daooh30','daoohr30','daooh90','daoohr90')],sd, na.rm = TRUE)
sapply(dt2[,c('daooh30','daoohr30','daooh90','daoohr90')],var, na.rm = TRUE)
```

-   There are some NA values in troponin and bnpval columns.

```{r}
# data distribution of daoohr30 and daoohr90
p1 <- ggplot(data = dt2, mapping = aes(x = daoohr30))+
  geom_histogram((aes(y = after_stat(count))))+
  scale_x_continuous(breaks = seq(0, 30, by = 10)) +
  labs(y = 'Frequency', x = 'DAOOH30')+
  theme_bw()

p2 <- ggplot(data = dt2, mapping = aes(x = daoohr90))+
  geom_histogram((aes(y = after_stat(count))))+
  scale_x_continuous(breaks = seq(0, 90, by = 15)) +
  labs(y = NULL,x = 'DAOOH90')+
  theme_bw()

p1 + p2

ggsave("hist.plot.png", plot = p1 + p2, width = 6, height = 4, dpi = 300)
```

# Multiple Imputation

reference: https://library.virginia.edu/data/articles/getting-started-with-multiple-imputation-in-r https://www.rdocumentation.org/packages/mice/versions/3.17.0/topics/mice

```{r}
# remove subjects whose daoohr are all NA values
dt2 <- dt2[!(is.na(dt2$daoohr30) & is.na(dt2$daoohr90)),]
# leave out the id variable
dt.im <- dt2[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval','treatgroup','daoohr30')]

# inpute missing values 
dt.imp <- mice(dt.im, m = 15, seed = 123) 
```

## check one complete dataset and use this for show the models

```{r}
df.comp1 <- complete(dt.imp, action = 1)  # extract first imputed dataset
head(df.comp1)
```

## pool the estimates together

```{r}
dt.long <- mice::complete(dt.imp, action="long", include = TRUE)
dt.long$daoohr30 <- dt2$daoohr30
dt.long$ndaoohr30 <- 30-dt2$daoohr30
dt.long$daoohr30.r <- ordered(dt2$daoohr30)
dt.long.mids <- as.mids(dt.long)
```

## imputed dt2 dataset

```{r}
dt2$troponinilow2 <- df.comp1$troponinilow2
dt2$bnpval <- df.comp1$bnpval
colSums(is.na(dt2))
head(dt2)
```

# T-test

```{r}
group <- factor(dt2$treatgroup, levels = c('SSU','Hospitalization'))
t.result <- t.test(daoohr30~group,data = dt2)
t.estimate <- t.result$estimate[1] - t.result$estimate[2]
t.ci.lw <- t.result$conf.int[1]
t.ci.up <- t.result$conf.int[2]
t.pval <- t.result$p.value
```

# Wilcoxon test

## Explore the arguments

reference: https://search.r-project.org/R/refmans/stats/html/wilcox.test.html

```{r}
U.result <- wilcox.test(daoohr30~treatgroup,data = dt2)
U.result

# if set exact = TRUE, get the warning 'cannot compute exact p-value with ties'
# if set exact = FALSE, the function use normal approximation for the distribution of U statistic
```

## Calculate the Probabilistic Index and 95% CI

The change in previous result may due to the NA values.

```{r}
grp1 <- dt2$daoohr30[dt2$treatgroup == 'SSU']
grp2 <- dt2$daoohr30[dt2$treatgroup == 'Hospitalization']
U <- U.result$statistic
n1 <- length(grp1)
n2 <- length(grp2)

# Probabilistic Index
wmw.estimate <- U / (n1 * n2)
wmw.p.val <- U.result$p.value
cat("Probabilistic Index (PI):", wmw.estimate, "\n")
# the probability that the daoohr30 of a random observation of the SSU group is larger or equal than that of a random observation of the Hospitalization  group
# if H0 holds, the probabilistic index = 1/2.
```

```{r}
# Bootstrap for Confidence Interval of PI
set.seed(123)
boot_pi <- replicate(1000, {
  sample1 <- sample(grp1, replace = TRUE)
  sample2 <- sample(grp2, replace = TRUE)
  U_boot <- wilcox.test(sample1,sample2)$statistic # still use correction here
  U_boot / (length(sample1) * length(sample2))
})

# 95% CI
CI <- quantile(boot_pi, probs = c(0.025, 0.975))
cat("95% Confidence Interval for PI:", CI, "\n")

wmw.ci.lw <- CI[1]
wmw.ci.up <- CI[2]
```

The reason for different result if change the position of sample 1 and sample 2: the function is comparing the the distribution of x(position 1) to the distribution of y(position 2).

```{r}
0.5 - 0.5128873
0.5 - 0.4840982 
```

```{r}
0.5 - 0.6817386
0.5 - 0.3147176
```

```{r}
# same p-values, different W statistic
wilcox.test(grp1,grp2)
wilcox.test(grp2,grp1)
```

# Modeling

## Linear Regression

```{r}
# model construction
model.linear <- lm(daoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup, data = dt2)
summary(model.linear)
```

-   sbp and bun show significant association with daoohr30

-   the effect of treatment group is not that significant

-   the intercept shows high significance: could it be that there is much variance in the response variable to be explained by other covariates?

-   adj R-squared is relatively small, the model is not fitted well

There seems no specific pattern in the residual vs fitted value plot, suggesting equal variance; however, the mean model may not be correct. qq-plot displays a departure from normality.

### Estimation result and Interpretation

The coefficient for treatgroupSSU is 1.332, which means given other covariates, subjects in the SSU treatment group are expected to have 1.337 more DAOOH than subjects in the Hospitalization group.

### pooled result

```{r}
dt.fit.linear <- with(dt.long.mids,
               lm(daoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup))

dt.pooled.linear <- pool(dt.fit.linear)
linear_pool_result <- summary(dt.pooled.linear)
linear_pool_result
```

### Diagnostic plots for linear regression

```{r}
fitted <- fitted(model.linear)
residuals <- residuals(model.linear)
plot_data <- data.frame(fitted = fitted, residuals = residuals)

# residuals vs fitted value
res.p1 <- ggplot(plot_data, aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "(a)", x = "Fitted Values", y = "Residuals")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))

# QQ-plot
qq.p1 <- ggplot(plot_data, aes(sample = residuals)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "(a)")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))

res.p1
qq.p1
```

### g-computation-pooled result

```{r}
m <- 15
g_m <- numeric(m)  # store g-computation result
pool_model <- dt.fit.linear

# for each model, make prediction based on the each dataset and calculate ATE
for (i in 1:m) {
  dt.i <- complete(dt.imp,i)
  dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
  predicted.i0 <- predict(pool_model$analyses[[i]],dt.i0)
  
  predicted.i0[predicted.i0 <0] <- 0
  predicted.i0[predicted.i0 >30] <- 30
  
  
  dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
  predicted.i1 <- predict(pool_model$analyses[[i]],dt.i1)
  
  predicted.i1[predicted.i1 <0] <- 0
  predicted.i1[predicted.i1 >30] <- 30
  
  g_m[i] <- mean(predicted.i1 - predicted.i0) # this is for linear regression
}

mean(g_m)
```

## Gamma Regression

```{r}
dt2$ndaoohr30 <- 30 - dt2$daoohr30
model.gamma <- glm(ndaoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup, data = dt2, family = Gamma(link = "log"))
summary(model.gamma)
```
### Diagnostic plots

Steps： 1.Simulate new response data from the fitted model for each observation.

2.For each observation, calculate the empirical cumulative density function for the simulated observations, which describes the possible values (and their probability) at the predictor combination of the observed value, assuming the fitted model is correct.

3.The residual is then defined as the value of the empirical density function at the value of the observed data, so a residual of 0 means that all simulated values are larger than the observed value, and a residual of 0.5 means half of the simulated values are larger than the observed value.

```{r}
simulationOutput <- simulateResiduals(fittedModel = model.gamma, plot = F)
# residuals(simulationOutput)
```

```{r}
plot(simulationOutput)
```

'containing the simulations and the scaled residuals, which can later be passed on to all other plots and test functions.' - can take out residuals

For a correctly specified model we would expect asymptotically a uniform distribution of the scaled residuals & uniformity in y direction if we plot against any predictor.

```{r}
res <- simulationOutput$scaledResiduals

fitted <- simulationOutput$fittedPredictedResponse # same as fitted(model.gamma)
plot_data <- data.frame(fitted = fitted, residuals = res)


# library(grid)
# library(gridGraphics)
plotResiduals(simulationOutput)

qq.p2 <- ggplot(plot_data, aes(sample = residuals)) +
  geom_qq(distribution = function(p) qunif(p, min = 0, max = 1)) +
  geom_abline(slope = 1, intercept = 0,  linetype = "dashed") +
  labs(title = "(b)")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))
qq.p2

plotQQunif(simulationOutput)


```

```{r}
png("dharma_base.png", width=480, height=480, res=100)

plotResiduals(simulationOutput,rank = T, main = "",quantreg = FALSE)
# a smooth spline around the mean

dev.off()

plotResiduals(simulationOutput,rank = T, main = "",quantreg = FALSE)
```

```{r}
plot_data <- plot_data[order(plot_data$fitted), ]
plot_data$rank_fitted <- rank(plot_data$fitted, ties.method = "average") / (nrow(plot_data) + 1)


res.p2 <- ggplot(plot_data, aes(x = rank_fitted, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(title = "(b)", x = "Rank Transformed Fitted Values", y = "Residuals")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))
res.p2
```

### Estimation and Interpretation

The model is:

$$
log(ndaoohr30)= \beta_0 + \beta_1hosp6mo+\beta_2sbp+\beta_3bun+\beta_4troponinilow2+\beta_5bnpval+\beta_6treatgroup
$$

The coefficient for treatgroupSSU is -0.2205, which means given other covariates, subjects in the SSU treatment group are expected to have a change in ndaoohr30 of exp(-0.2205×1) = 0.802 times compared to subjects in the Hospitalization group.(within the context of truncated sample)

### pooled result

```{r}
dt.fit.gamma <- with(dt.long.mids,
               glm(ndaoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup, family = Gamma(link = "log")))

dt.pooled.gamma <- pool(dt.fit.gamma)
gamma_pool_result <- summary(dt.pooled.gamma)
gamma_pool_result
```

### g-computation-pooled result

```{r}
m <- 15
g_m <- numeric(m)  # store g-computation result
pool_model <- dt.fit.gamma

# for each model, make prediction based on the each dataset and calculate ATE
for (i in 1:m) {
  dt.i <- complete(dt.imp,i)
  dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
  predicted.i0 <- predict(pool_model$analyses[[i]],dt.i0)
  
  predicted.i0[predicted.i0 > log(30)] <- log(30)
  
  dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
  predicted.i1 <- predict(pool_model$analyses[[i]],dt.i1)
  
  predicted.i1[predicted.i1 > log(30)] <- log(30)
  
  g_m[i] <- mean(30 - exp(predicted.i1) - (30 - exp(predicted.i0))) # this is for gamma regression, should first exponentiate and then 30 - to get the daoohr30
}

mean(g_m)
```

## Ordinal Regression (RMS package)

```{r}
library(rms)
```

reference: https://cran.r-project.org/web/packages/rms/rms.pdf

```{r}
hosp6mo <- dt2$hosp6mo; bun <- dt2$bun; troponinilow2 <- dt2$troponinilow2; bnpval <- dt2$bnpval; treatgroup <- dt2$treatgroup; sbp <- dt2$sbp; daoohr30 <- dt2$daoohr30

ddist <- datadist(hosp6mo, bun, troponinilow2, bnpval, treatgroup, sbp,daoohr30)
options(datadist = ddist)

model.ord.r <- orm(daoohr30 ~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, x = TRUE, y = TRUE)

summary(model.ord.r)
```

```{r}
predict(model.ord.r,type = 'mean')
# the estimated mean Y by summing values of Y multiplied by the estimated Prob(Y= j).
```

### Diagnostic plots

```{r}
fitted <- predict(model.ord.r,type = 'mean')
```

```{r}
# PResiduals
# https://cran.r-project.org/src/contrib/Archive/PResiduals/?C=M;O=A
# install.packages("PResiduals_1.0-1.tar", repos = NULL, type = "source")
```

```{r}
library("PResiduals")
```

```{r}
# Residual-by-predictor plots with PSRs from proportional odds model
PSR.orm <- presid(model.ord.r)

plot(fitted, PSR.orm, 
               xlab = "The Fitted Value", ylab = "PSRs", 
               col = gray(.6),
               main = "Residual-by-Fitted Plot", 
               cex = 0.3, cex.main = 0.8)
lines(lowess(fitted, PSR.orm), lwd = 3)
abline(h = 0, lty = 2)


# residuals vs fitted value
plot_data <- data.frame(fitted = fitted, residuals = PSR.orm)

res.p3 <- ggplot(plot_data, aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "(c)", x = "Fitted Values", y = "Residuals")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))
res.p3

```

F(Y) is the probability integral transformation and it is uniformly distributed from 0 to 1. Therefore, if the PSR is from the properly-specified model, it is a re-scaling of the probability integral transformation and it is approximately uniformly distributed from−1 to 1 with expectation 0 and constant variance 1/3. A quantile-quantile (QQ) plot of PSRs versus the theoretical quantiles of the uniform distribution can be used to assess the overall model fit.

```{r}
# QQ-plot
qqplot(qunif(ppoints(length(PSR.orm)), -1, 1), PSR.orm, 
  main = " QQ-plots of probability-scale residuals (PSRs)", cex = 0.4, xlab = "Quantiles of Uniform Distribution",
  ylab = "Quantiles of PSRs") 
abline(a = 0, b = 1, lty = 2)
grid()

qq.p3 <- ggplot(plot_data, aes(sample = PSR.orm)) +
  geom_qq(distribution = function(p) qunif(p, min = -1, max = 1)) +
  geom_abline(slope = 1, intercept = 0,  linetype = "dashed") +
  labs(title = "(c)")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))
qq.p3

```

### pooled result

\*\*\* compare aregImpute and mice: https://stats.stackexchange.com/questions/295977/how-to-use-restricted-cubic-splines-with-the-r-mice-imputation-package

Like mice, aregImpute uses chained equations. Unlike mice, it uses bootstrap draws instead of approximate (assuming multivariate normality) Bayesian posterior draws.

```{r}
im <- aregImpute(~ daoohr30+hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, 
                 data = dt.im,
                 n.impute = 15)
dt.pooled.orm <- fit.mult.impute(daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp , xtrans = im, data = dt.im, fitter = orm)
orm_pool_result <- summary(dt.pooled.orm)
orm_pool_result
```

### mice + polr result

```{r}
# mice + polr
dt.fit.ord <- with(dt.long.mids,
               polr(daoohr30.r~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, Hess=TRUE))

dt.pooled.ord <- pool(dt.fit.ord)
ord_pool_result <- summary(dt.pooled.ord)
ord_pool_result
```

### g-computation pooled result

```{r}
m <- 15
g_m <- numeric(m)  


# for each model, make prediction based on the each dataset and calculate ATE
for (i in 1:m) {
  dt.i <- impute.transcan(im, data=dt.im, imputation=i, list.out=TRUE, pr=FALSE, check=FALSE)

# convert the list to the database
  dt.i <- as.data.frame(dt.i, stringsAsFactors = FALSE)
  ddist <- datadist(dt.i)
  options(datadist = ddist)
  
  model.i <- orm(daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, x = TRUE, y = TRUE, data = dt.i)
  
  dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
  
  predicted.i0 <- predict(model.i,dt.i0,type = 'mean')
  
  dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
  predicted.i1 <- predict(model.i,dt.i1,type = 'mean')
  
  
  g_m[i] <- mean(predicted.i1 - predicted.i0) 
}

mean(g_m)
```

## Probabilistic Index Model

reference: 1. https://cran.r-project.org/web/packages/pim/vignettes/pim.pdf 2. De Schryver M, De Neve J. A tutorial on probabilistic index models: Regression models for the effect size P(Y1 \< Y2). Psychol Methods. 2019 Aug;24(4):403-418. doi: 10.1037/met0000194. Epub 2018 Sep 27. PMID: 30265047.(Wilcoxon–Mann–Whitney (WMW) test, t test, PIM)

```{r}
model.pim <- pim(daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, data = dt2)
summary(model.pim)
```

### Estimation and Interpretation

```{r}
exp(model.pim@coef['treatgroupSSU'])/(1 + exp(model.pim@coef['treatgroupSSU']))
```

The expit coefficient for treatgroupSSU is 0.60, which means given other covariates, the probability that subjects in the SSU treatment group have a higher daoohr30 compared to subjects in the Hospitalization group is estimated by 0.60.

### pooled result manually based on Rubin’s Rules

```{r}
pim_pool <- function(imp, formula, variable = "treatgroupSSU") {
  m <- imp$m
  q_m <- numeric(m)
  u_m <- numeric(m)
  
  for (i in 1:m) {
    dt.i <- complete(imp, i)
    formula_obj <- as.formula(formula)
    
    model <- pim(formula_obj, data = dt.i)

    coef <- model@coef[variable]
    se <- summary(model)@se[variable]
    
    q_m[i] <- coef
    u_m[i] <- se^2
    
  }
  
  # Rubin’s Rules pooling
  Q_bar <- mean(q_m)
  U_bar <- mean(u_m)
  B <- var(q_m)
  T_var <- U_bar + (1 + 1/m) * B
  se_pooled <- sqrt(T_var)
  t_value <- Q_bar / se_pooled
  df <- (m - 1) * (1 + U_bar / ((1 + 1/m) * B))^2
  p_value <- 2 * pt(-abs(t_value), df)
  

  result <- list(
    estimate = Q_bar,
    std.error = se_pooled,
    t.value = t_value,
    df = df,
    p.value = p_value
  )
  
  return(result)
}

```

```{r}
pim_pool_result <- pim_pool(imp  = dt.imp, formula = 'daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp')
pim_pool_result
```

### compare with WMW test

```{r}
# same p-val as Wilcoxon–Mann–Whitney test
summary(pim(daoohr30~ treatgroup, data = dt2))
```

# G-computation CI

```{r}
#| echo: false
#| results: hide
library(Hmisc)
library(ggplot2)
library(rms)
library(mice)
library(boot)
library(MASS)
library(gt)
```

```{r}
# set the seed
set.seed(123)
```

## Functions

```{r}
bootstrap_data <- function(data) {
  
  n <- nrow(data)
  boot_data <- data[sample(1:n, size = n, replace = TRUE), ]
  
  return(boot_data)
  
}

glm_g_comp_one_sample_linear <- function(sample_imp_data, pool_model) {
  
  m <- m
  g_m <- numeric(m)  # store g-computation result
  
  # for each model, make prediction based on the each dataset and calculate ATE
  for (i in 1:m) {
    dt.i <- complete(sample_imp_data,i)
    dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
    predicted.i0 <- predict(pool_model$analyses[[i]],dt.i0)
    
    predicted.i0[predicted.i0 <0] <- 0
    predicted.i0[predicted.i0 >30] <- 30
    
    dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
    predicted.i1 <- predict(pool_model$analyses[[i]],dt.i1)
    
    predicted.i1[predicted.i1 <0] <- 0
    predicted.i1[predicted.i1 >30] <- 30
    
    # check values
    if (any(predicted.i0 < 0)) {
      print('negative values')
    } else if(any(predicted.i0 > 30)) {
      print('>30 values')
    } 
    
    if (any(predicted.i1 < 0)) {
      print('negative values1')
    } else if(any(predicted.i1 > 30)) {
      print('>30 values1')
    }
    
    g_m[i] <- mean(predicted.i1 - predicted.i0) # this is for linear regression
  }
  
  return(g_m)
  
}

glm_g_comp_one_sample_gamma <- function(sample_imp_data, pool_model) {
  
  m <- m
  g_m <- numeric(m)  # store g-computation result
  
  # for each model, make prediction based on the each dataset and calculate ATE
  for (i in 1:m) {
    dt.i <- complete(sample_imp_data,i)
    dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
    predicted.i0 <- predict(pool_model$analyses[[i]],dt.i0)
    
    predicted.i0[predicted.i0 > log(30)] <- log(30)
    
    dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
    
    predicted.i1 <- predict(pool_model$analyses[[i]],dt.i1)
    
    predicted.i1[predicted.i1 > log(30)] <- log(30)
    
    # check values
    if (any(predicted.i0 > log(30))) {
      print('>30 values')
    }
    
    if (any(predicted.i1 > log(30))) {
      print('>30 values1')
    }
    
    
    g_m[i] <- mean(30 - exp(predicted.i1) - (30 - exp(predicted.i0))) # this is for gamma regression, should first exponentiate and then 30 - to get the daoohr30
  }
  
  return(g_m)
  
}

glm_g_comp_one_sample_ordinal <- function(sample_imp_data, pool_model) {
  
  m <- m
  g_m <- numeric(m)  # store g-computation result
  
  # for each model, make prediction based on the each dataset and calculate ATE
  for (i in 1:m) {
    dt.i <- impute.transcan(pool_model, data=sample_imp_data, imputation=i, list.out=TRUE, pr=FALSE, check=FALSE)

  # convert the list to the database
    dt.i <- as.data.frame(dt.i, stringsAsFactors = FALSE)
    ddist <- datadist(dt.i)
    options(datadist = ddist)
    
    model.i <- orm(daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, x = TRUE, y = TRUE, data = dt.i)
    
    dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
    
    predicted.i0 <- predict(model.i,dt.i0,type = 'mean')
    
    dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
    predicted.i1 <- predict(model.i,dt.i1,type = 'mean')
    
    
    g_m[i] <- mean(predicted.i1 - predicted.i0)
    
  }
  
  return(g_m)
  
}


# store the result of all bootstrap result (n times of bootstrap) at original scale
g_comp_glm_result_linear <- function(n, m, family, data){
  
  result <- matrix(NA, nrow = n, ncol = m) #store the original scale result
  
  for (j in 1:n) {
    # bootstrap
    sample_data <- bootstrap_data(data)
    
    # multiple imputation
    suppressMessages(suppressWarnings(
      invisible(capture.output({
        sample_imp_data <- mice(sample_data, m = m, quiet = TRUE)
        }))
      ))
    
    sample_imp_long <- complete(sample_imp_data, action="long", include = TRUE)
    sample_imp_long$daoohr30 <- sample_data$daoohr30
    sample.mids <- as.mids(sample_imp_long)
    
    # fit the model
    pool_model <- with(sample.mids,
                       glm(formula = daoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup,
                           family = family))
    result[j, ] <- glm_g_comp_one_sample_linear(sample_imp_data,pool_model)
    
  }
  
  result_df <- as.data.frame(result)
  effect_of_interest <- rowMeans(result, na.rm = TRUE)
  
  return(effect_of_interest)
  
}


g_comp_glm_result_gamma <- function(n, m, family, data){
  
  result <- matrix(NA, nrow = n, ncol = m) #store the original scale result
  
  for (j in 1:n) {
    # bootstrap
    sample_data <- bootstrap_data(data)
    
    # multiple imputation
    suppressMessages(suppressWarnings(
      invisible(capture.output({
        sample_imp_data <- mice(sample_data, m = m, quiet = TRUE)
        }))
      ))
    
    sample_imp_long <- complete(sample_imp_data, action="long", include = TRUE)
    sample_imp_long$ndaoohr30 <- 30-sample_data$daoohr30
    sample.mids <- as.mids(sample_imp_long)
    
    # fit the model
    pool_model <- with(sample.mids,
                       glm(formula = ndaoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup,
                           family = family))
    result[j, ] <- glm_g_comp_one_sample_gamma(sample_imp_data,pool_model)
    
  }
  
  result_df <- as.data.frame(result)
  effect_of_interest <- rowMeans(result, na.rm = TRUE)
  
  return(effect_of_interest)
  
} # get an n row vector here, in daoohr30 scale

g_comp_glm_result_ordinal <- function(n, m, data){
  
  result <- matrix(NA, nrow = n, ncol = m) #store the original scale result
  
  for (j in 1:n) {
    # bootstrap
    sample_data <- bootstrap_data(data)
    
    # multiple imputation
    suppressMessages(suppressWarnings(
      invisible(capture.output({
        im <- aregImpute(~ daoohr30+hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp,
                     n.impute=m,
                     data = sample_data)
        }))
      ))
    
    
    result[j, ] <- glm_g_comp_one_sample_ordinal(sample_imp_data = sample_data,pool_model=im)
    
  }
  
  result_df <- as.data.frame(result)
  effect_of_interest <- rowMeans(result, na.rm = TRUE)
  
  return(effect_of_interest)
  
}

```

## Arguments

```{r}
m = 15
n = 500
```

## Prepare for Data

```{r}
# new analysis data
dt2 <- read.csv('analytic_data.csv')

# transform data
dt2$daooh30 <- as.numeric(dt2$daooh30)
dt2$daoohr30 <- as.numeric(dt2$daoohr30)
dt2$daooh90 <- as.numeric(dt2$daooh90)
dt2$daoohr90 <- as.numeric(dt2$daoohr90)

dt2$hosp6mo <- as.factor(dt2$hosp6mo)
dt2$troponinilow2 <- as.factor(dt2$troponinilow2)
```

```{r}
# remove subjects whose daoohr are all NA values
dt2 <- dt2[!(is.na(dt2$daoohr30) & is.na(dt2$daoohr90)),]

# leave out the id variable
dt2 <- dt2[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval','treatgroup','daoohr30')]
```

## CI calculation for linear regression

```{r}
# n, m, family, formula, data

linear_result <- g_comp_glm_result_linear(n = n,
                                          m = m,
                                          family = gaussian(link = "identity"), 
                                          data = dt2)

g_comp_linear <- quantile(linear_result,c(0.025,0.5,0.975))
g_comp_linear
```

## CI calculation for gamma regression

```{r}
# n, m, family, formula, data
gamma_result <- g_comp_glm_result_gamma(n = n,
                                        m = m,
                                        family = Gamma(link = "log"), 
                                        data = dt2)

g_comp_gamma <- quantile(gamma_result,c(0.025,0.5,0.975))
g_comp_gamma
```

## CI calculation for ordinal regression

```{r}
# n, m, family, formula, data

ordinal_result <- g_comp_glm_result_ordinal(n = n,
                                  m = m, 
                                  data = dt2)

g_comp_ord <- quantile(ordinal_result,c(0.025,0.5,0.975))
g_comp_ord
```

# Summary table for result

how the plot will look like: model,estimand, estimate, CI and p value

separate the mean model and multiplicative model by blank space

mean model: ?t-test; linear regression; probabilistic model

multiplicative model: gamma regresison; ordinal regression; ?WMW test; pim model;

## Result Table 1: estimates, CI, p-val, estimand

```{r}
# collect the estimates

trt30r.estimate <- c(NA,
                     t.estimate,
                     linear_pool_result[7,'estimate'], # result of treatgroupSSU
                     NA,
                     NA,
                     exp(gamma_pool_result[7,'estimate']),
                     exp(orm_pool_result['treatgroup - SSU:Hospitalization','Effect']),
                     NA,
                     NA,
                     exp(wmw.estimate)/(1 + exp(wmw.estimate)),
                     exp(pim_pool_result$estimate)/(1 + exp(pim_pool_result$estimate))
                     )

trt30r.estimand <- c(NA,
                     'Mean',
                     'Mean',
                     NA,
                     NA,
                     'Geometreic Mean',
                     'Odds Ratio',
                     NA,
                     NA,
                     'Probabilistic Index',
                     'Probabilistic Index'
                     )

trt30r.lw <- c(NA,
               t.ci.lw,
               linear_pool_result[7,'estimate']-1.96*linear_pool_result[7,'std.error'],
               NA,
               NA,
               exp(gamma_pool_result[7,'estimate']-1.96*gamma_pool_result[7,'std.error']),
               exp(orm_pool_result['treatgroup - SSU:Hospitalization','Lower 0.95']),
               NA,
               NA,
               wmw.ci.lw,
               exp(pim_pool_result$estimate-1.96*pim_pool_result$std.error)/(1+exp(pim_pool_result$estimate-1.96*pim_pool_result$std.error))
               )

trt30r.up <- c(NA,
               t.ci.up,
               linear_pool_result[7,'estimate']+1.96*linear_pool_result[7,'std.error'],
               NA,
               NA,
               exp(gamma_pool_result[7,'estimate']+1.96*gamma_pool_result[7,'std.error']),
               exp(orm_pool_result['treatgroup - SSU:Hospitalization','Upper 0.95']),
               NA,
               NA,
               wmw.ci.up,
               exp(pim_pool_result$estimate+1.96*pim_pool_result$std.error)/(1+exp(pim_pool_result$estimate+1.96*pim_pool_result$std.error))
               )

# p_val for ordinal regression
L <- orm_pool_result['treatgroup - SSU:Hospitalization','Lower 0.95']
U <- orm_pool_result['treatgroup - SSU:Hospitalization','Upper 0.95']
SE <- orm_pool_result['treatgroup - SSU:Hospitalization','S.E.']
Est <- orm_pool_result['treatgroup - SSU:Hospitalization','Effect']
z <- Est/SE
ord_p_val <- 2 * (1 - pnorm(abs(z)))


trt30r.pval <- c(NA,
                 t.pval,
                 linear_pool_result[7,'p.value'],
                 NA,
                 NA,
                 gamma_pool_result[7,'p.value'],
                 ord_p_val, # use the formula
                 NA,
                 NA,
                 wmw.p.val,
                 pim_pool_result$p.value
                 )


trt30r.label <- c("Additive Model","T-test","Linear",'',
                  "Multiplicative Model","Gamma","Ordinal",'',
                  "Probabilistic Model","WMW-test","PIM")

# prepare the dataframe for plot
(dt.trt30r <- data.frame(subgroup = trt30r.label,
                         estimate = round(trt30r.estimate,3),
                         estimand = trt30r.estimand,
                         ci.low = round(trt30r.lw,3),
                         ci.up = round(trt30r.up,3),
                         pval = round(trt30r.pval,3)))
```

```{r}
# add the blank column to display CI

dt.trt30r$" " <- paste(rep(" ", 20), collapse = " ")

# Create a confidence interval column to display
dt.trt30r$`95% CI` <- ifelse((is.na(dt.trt30r$ci.low)), "",
                             sprintf("(%.2f to %.2f)",
                                     dt.trt30r$ci.low, dt.trt30r$ci.up))

dt.trt30r$estimate <- ifelse(is.na(dt.trt30r$estimate),"",dt.trt30r$estimate)
dt.trt30r$ci.low <- ifelse(is.na(dt.trt30r$ci.low),"",dt.trt30r$ci.low)
dt.trt30r$ci.up <- ifelse(is.na(dt.trt30r$ci.up),"",dt.trt30r$ci.up)
dt.trt30r$pval <- ifelse(is.na(dt.trt30r$pval),"",dt.trt30r$pval)
dt.trt30r$estimand <- ifelse(is.na(dt.trt30r$estimand),"",dt.trt30r$estimand)

dt.trt30r$subgroup <- ifelse(is.na(dt.trt30r$estimate), dt.trt30r$subgroup,paste0("   ", dt.trt30r$subgroup))
dt.trt30r[c('subgroup','estimate','estimand','pval','95% CI')]
```

```{r}
dt.trt30r[c('subgroup','estimate','estimand','pval','95% CI')] %>%
  gt() %>%
  tab_header(title = "Table 1") %>%
  fmt_number(columns = c("pval"), decimals = 2) %>%
  cols_label(
    subgroup = "Subgroup",
    estimate = "Estimate",
    estimand = "Estimand",
    pval = "P-value",
    '95% CI' = "95%CI"
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(
      rows = subgroup %in% c("Additive Model", "Multiplicative Model"),
      columns = everything()
    )
  )
```


## Result Table 2: g-computation result: linear, gamma, ord

```{r}
# collect the estimates

g.comp.estimate <- c(NA,
                     g_comp_linear[2],
                     NA,
                     NA,
                     g_comp_gamma[2],
                     g_comp_ord[2]
                     )


g.comp.lw <- c(NA,
               g_comp_linear[1],
               NA,
               NA,
               g_comp_gamma[1],
               g_comp_ord[1]
               )

g.comp.up <- c(NA,
               g_comp_linear[3],
               NA,
               NA,
               g_comp_gamma[3],
               g_comp_ord[3]
               )



g.comp.label <- c("Additive Model","Linear",'',
                  "Multiplicative Model","Gamma","Ordinal")

# prepare the dataframe for plot
(dt.g.comp <- data.frame(subgroup = g.comp.label,
                         estimate = round(g.comp.estimate,3),
                         ci.low = round(g.comp.lw,3),
                         ci.up = round(g.comp.up,3)))
```

```{r}

dt.g.comp$" " <- paste(rep(" ", 20), collapse = " ")

# Create a confidence interval column to display
dt.g.comp$`95% CI` <- ifelse((is.na(dt.g.comp$ci.low)), "",
                             sprintf("(%.2f to %.2f)",
                                     dt.g.comp$ci.low, dt.g.comp$ci.up))

dt.g.comp$estimate <- ifelse(is.na(dt.g.comp$estimate),"",dt.g.comp$estimate)

dt.g.comp$subgroup <- ifelse(is.na(dt.g.comp$estimate), dt.g.comp$subgroup,paste0("   ", dt.g.comp$subgroup))
dt.g.comp[c('subgroup','estimate','95% CI')]
```

```{r}
dt.g.comp[c('subgroup','estimate','95% CI')] %>%
  gt() %>%
  tab_header(title = "Table 2") %>%
  cols_label(
    subgroup = "Subgroup",
    estimate = "Estimate",
    '95% CI' = "95%CI"
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(
      rows = subgroup %in% c("Additive Model", "Multiplicative Model"),
      columns = everything()
    )
  )
```

## Diagnostic Plots

```{r}
res.p1.n <- res.p1 + labs(y = 'Residuals',x = 'DAOOH') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
res.p2.n <- res.p2 + labs(y = 'Residuals',x = 'DAOOH(Rank transformed)') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
res.p3.n <- res.p3 + labs(y = 'PResiduals',x = 'DAOOH') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))

res.plot <- ggarrange(res.p1.n, res.p2.n,res.p3.n,
                       ncol = 3, nrow = 1,
                       align = "hv",
                       widths = c(1, 1),
                       common.legend = FALSE)
res.plot <- annotate_figure(res.plot)
res.plot
ggsave("res.plot.png", plot = res.plot, width = 10, height = 3, dpi = 300)
```

```{r}
res.p1.n <- res.p1 + labs(y = NULL, x = 'Quantiles of Normal Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5), axis.title.x = element_text(size = 6))
res.p2.n <- res.p2 + labs(y = NULL, x = 'Quantiles of Normal Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5), axis.title.x = element_text(size = 6))
res.p3.n <- res.p3 + labs(y = NULL, x = 'Quantiles of Uniform Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5), axis.title.x = element_text(size = 6))

res.plot <- ggarrange(res.p1.n, res.p2.n,res.p3.n,
                       ncol = 1, nrow = 3,
                       align = "hv",
                       widths = c(1, 1),
                       common.legend = FALSE)
res.plot <- annotate_figure(res.plot)
res.plot

ggsave("res.plot_v.png", res.plot, width = 10, height = 3, dpi = 300)
```

```{r}
qq.p1.n <- qq.p1 + labs(y = NULL, x = 'Quantiles of Normal Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
qq.p2.n <- qq.p2 + labs(y = NULL, x = 'Quantiles of Uniform Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
qq.p3.n <- qq.p3 + labs(y = NULL, x = 'Quantiles of Uniform Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
qq.plot <- ggarrange(qq.p1.n, qq.p2.n,qq.p3.n,
                       ncol = 3, nrow = 1,
                       align = "hv",
                       widths = c(1, 1),
                       common.legend = FALSE)
qq.plot <- annotate_figure(qq.plot,
                         left   = text_grob("Sample Quantiles", rot = 90, size = 10),
                         bottom = text_grob("Theoretical Quantiles", size = 10))
qq.plot
ggsave("qq.plot.png", plot = qq.plot, width = 10, height = 3, dpi = 600)
```
