---
title: "SSU_week_17"
format: html
editor: visual
---

```{r}
set.seed(123)
```

```{r}
#| echo: false
#| results: hide
library(Hmisc)
library(ggplot2)
library(rms)
library(mice)
library(boot)
library(MASS)
library(pim)
library(patchwork)
library(ggpubr)
library(DHARMa)
```

# Check Analysis Data

```{r}
dt2 <- read.csv('analytic_data.csv')

# transform data
dt2$daooh30 <- as.numeric(dt2$daooh30)
dt2$daoohr30 <- as.numeric(dt2$daoohr30)
dt2$daooh90 <- as.numeric(dt2$daooh90)
dt2$daoohr90 <- as.numeric(dt2$daoohr90)

dt2$hosp6mo <- as.factor(dt2$hosp6mo)
dt2$troponinilow2 <- as.factor(dt2$troponinilow2)
```

```{r}
colnames(dt2)
```

```{r}
# data summary
table2 <- summaryM(daooh30+daooh90+daoohr30+daoohr90+hosp6mo+sbp+bun+troponinilow2+bnpval ~treatgroup,data = dt2)
html(table2)

sapply(dt2[,c('daooh30','daoohr30','daooh90','daoohr90')],sd, na.rm = TRUE)
sapply(dt2[,c('daooh30','daoohr30','daooh90','daoohr90')],var, na.rm = TRUE)
```

-   There are some NA values in troponin and bnpval columns.

```{r}
# data distribution of daoohr30 and daoohr90
p1 <- ggplot(data = dt2, mapping = aes(x = daoohr30))+
  geom_histogram((aes(y = after_stat(count))))+
  scale_x_continuous(breaks = seq(0, 30, by = 10)) +
  labs(y = 'Frequency', x = 'DAOOH30')+
  theme_bw()

p2 <- ggplot(data = dt2, mapping = aes(x = daoohr90))+
  geom_histogram((aes(y = after_stat(count))))+
  scale_x_continuous(breaks = seq(0, 90, by = 15)) +
  labs(y = NULL,x = 'DAOOH90')+
  theme_bw()

p1 + p2

ggsave("hist.plot.png", plot = p1 + p2, width = 6, height = 4, dpi = 300)
```

# Multiple Imputation

```{r}
# remove subjects whose daoohr are all NA values
dt2 <- dt2[!(is.na(dt2$daoohr30) & is.na(dt2$daoohr90)),]
dt.im <- dt2[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval','treatgroup','daoohr30')]
dt.imp <- mice(dt.im, m = 15, seed = 123) 
```

## one complete dataset

```{r}
df.comp1 <- complete(dt.imp, action = 1)
head(df.comp1)
```

## pool the estimates together

```{r}
dt.long <- mice::complete(dt.imp, action="long", include = TRUE)
dt.long$daoohr30 <- dt2$daoohr30
dt.long$ndaoohr30 <- 30-dt2$daoohr30
dt.long$daoohr30.r <- ordered(dt2$daoohr30)
dt.long.mids <- as.mids(dt.long)
```

## imputed dt2 dataset

```{r}
dt2$troponinilow2 <- df.comp1$troponinilow2
dt2$bnpval <- df.comp1$bnpval
colSums(is.na(dt2))
head(dt2)
```

# T-test

```{r}
group <- factor(dt2$treatgroup, levels = c('SSU','Hospitalization'))
t.result <- t.test(daoohr30~group,data = dt2)
t.estimate <- t.result$estimate[1] - t.result$estimate[2]
t.ci.lw <- t.result$conf.int[1]
t.ci.up <- t.result$conf.int[2]
t.pval <- t.result$p.value
```

# Wilcoxon test

```{r}
U.result <- wilcox.test(daoohr30~treatgroup,data = dt2)
U.result
```

## Calculate the Probabilistic Index and 95% CI

```{r}
grp1 <- dt2$daoohr30[dt2$treatgroup == 'SSU']
grp2 <- dt2$daoohr30[dt2$treatgroup == 'Hospitalization']
U <- U.result$statistic
n1 <- length(grp1)
n2 <- length(grp2)

# Probabilistic Index
wmw.estimate <- U / (n1 * n2)
wmw.p.val <- U.result$p.value
cat("Probabilistic Index (PI):", wmw.estimate, "\n")
```

```{r}
# Bootstrap for Confidence Interval of PI
set.seed(123)
boot_pi <- replicate(1000, {
  sample1 <- sample(grp1, replace = TRUE)
  sample2 <- sample(grp2, replace = TRUE)
  U_boot <- wilcox.test(sample1,sample2)$statistic
  U_boot / (length(sample1) * length(sample2))
})

# 95% CI
CI <- quantile(boot_pi, probs = c(0.025, 0.975))
cat("95% Confidence Interval for PI:", CI, "\n")

wmw.ci.lw <- CI[1]
wmw.ci.up <- CI[2]
```

```{r}
0.5 - 0.5128873
0.5 - 0.4840982 
```

```{r}
0.5 - 0.6817386
0.5 - 0.3147176
```

```{r}
# same p-values
wilcox.test(grp1,grp2)
wilcox.test(grp2,grp1)
```

# Modeling

## Linear Regression

```{r}
# model construction
model.linear <- lm(daoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup, data = dt2)
summary(model.linear)
```

### Estimation result and Interpretation

The coefficient for treatgroupSSU is 1.40, which means given other covariates, subjects in the SSU treatment group are expected to have 1.40 more DAOOH than subjects in the Hospitalization group.

### pooled result

```{r}
dt.fit.linear <- with(dt.long.mids,
               lm(daoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup))

dt.pooled.linear <- pool(dt.fit.linear)
linear_pool_result <- summary(dt.pooled.linear)
linear_pool_result
```

### Diagnostic plots for linear regression

```{r}
fitted <- fitted(model.linear)
residuals <- residuals(model.linear)
plot_data <- data.frame(fitted = fitted, residuals = residuals)

# residuals vs fitted value
res.p1 <- ggplot(plot_data, aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "(a)", x = "Fitted Values", y = "Residuals")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))

# QQ-plot
qq.p1 <- ggplot(plot_data, aes(sample = residuals)) +
  geom_qq() +
  geom_qq_line() +
  labs(title = "(a)")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))

res.p1
qq.p1
```

### g-computation-pooled result

```{r}
m <- 15
g_m <- numeric(m) 
pool_model <- dt.fit.linear

# for each model, make prediction based on the each dataset and calculate ATE
for (i in 1:m) {
  dt.i <- complete(dt.imp,i)
  dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
  predicted.i0 <- predict(pool_model$analyses[[i]],dt.i0)
  
  predicted.i0[predicted.i0 <0] <- 0
  predicted.i0[predicted.i0 >30] <- 30
  
  
  dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
  predicted.i1 <- predict(pool_model$analyses[[i]],dt.i1)
  
  predicted.i1[predicted.i1 <0] <- 0
  predicted.i1[predicted.i1 >30] <- 30
  
  g_m[i] <- mean(predicted.i1 - predicted.i0) # this is for linear regression
}

mean(g_m)
```

## Gamma Regression

```{r}
dt2$ndaoohr30 <- 30 - dt2$daoohr30
model.gamma <- glm(ndaoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup, data = dt2, family = Gamma(link = "log"))
summary(model.gamma)
```

### Diagnostic plots

```{r}
simulationOutput <- simulateResiduals(fittedModel = model.gamma, plot = F)
```

```{r}
plot(simulationOutput)
```

For a correctly specified model we would expect asymptotically a uniform distribution of the scaled residuals & uniformity in y direction if we plot against any predictor.

```{r}
res <- simulationOutput$scaledResiduals

fitted <- simulationOutput$fittedPredictedResponse # same as fitted(model.gamma)
plot_data <- data.frame(fitted = fitted, residuals = res)

plotResiduals(simulationOutput)

qq.p2 <- ggplot(plot_data, aes(sample = residuals)) +
  geom_qq(distribution = function(p) qunif(p, min = 0, max = 1)) +
  geom_abline(slope = 1, intercept = 0,  linetype = "dashed") +
  labs(title = "(b)")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))
qq.p2

plotQQunif(simulationOutput)
```


```{r}
plot_data <- plot_data[order(plot_data$fitted), ]
plot_data$rank_fitted <- rank(plot_data$fitted, ties.method = "average") / (nrow(plot_data) + 1)


res.p2 <- ggplot(plot_data, aes(x = rank_fitted, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(title = "(b)", x = "Rank Transformed Fitted Values", y = "Residuals")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))
res.p2
```

### Estimation and Interpretation

The model is:

$$
log(ndaoohr30)= \beta_0 + \beta_1hosp6mo+\beta_2sbp+\beta_3bun+\beta_4troponinilow2+\beta_5bnpval+\beta_6treatgroup
$$

The coefficient for treatgroupSSU is -0.22, which means given other covariates, subjects in the SSU treatment group are expected to have a change in ndaoohr30 of exp(-0.22×1) = 0.80 times compared to subjects in the Hospitalization group.

### pooled result

```{r}
dt.fit.gamma <- with(dt.long.mids,
               glm(ndaoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup, family = Gamma(link = "log")))

dt.pooled.gamma <- pool(dt.fit.gamma)
gamma_pool_result <- summary(dt.pooled.gamma)
gamma_pool_result
```

### g-computation-pooled result

```{r}
m <- 15
g_m <- numeric(m)  # store g-computation result
pool_model <- dt.fit.gamma

# for each model, make prediction based on the each dataset and calculate ATE
for (i in 1:m) {
  dt.i <- complete(dt.imp,i)
  dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
  predicted.i0 <- predict(pool_model$analyses[[i]],dt.i0)
  
  predicted.i0[predicted.i0 > log(30)] <- log(30)
  
  dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
  predicted.i1 <- predict(pool_model$analyses[[i]],dt.i1)
  
  predicted.i1[predicted.i1 > log(30)] <- log(30)
  
  g_m[i] <- mean(30 - exp(predicted.i1) - (30 - exp(predicted.i0)))
  
}

mean(g_m)
```

## Ordinal Regression (RMS package)

```{r}
library(rms)
```

```{r}
hosp6mo <- dt2$hosp6mo; bun <- dt2$bun; troponinilow2 <- dt2$troponinilow2; bnpval <- dt2$bnpval; treatgroup <- dt2$treatgroup; sbp <- dt2$sbp; daoohr30 <- dt2$daoohr30

ddist <- datadist(hosp6mo, bun, troponinilow2, bnpval, treatgroup, sbp,daoohr30)
options(datadist = ddist)

model.ord.r <- orm(daoohr30 ~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, x = TRUE, y = TRUE)

summary(model.ord.r)
```

```{r}
predict(model.ord.r,type = 'mean')
```

### Diagnostic plots

```{r}
fitted <- predict(model.ord.r,type = 'mean')
```

```{r}
library("PResiduals")
```

```{r}
# Residual-by-predictor plots with PSRs from proportional odds model
PSR.orm <- presid(model.ord.r)

plot(fitted, PSR.orm, 
               xlab = "The Fitted Value", ylab = "PSRs", 
               col = gray(.6),
               main = "Residual-by-Fitted Plot", 
               cex = 0.3, cex.main = 0.8)
lines(lowess(fitted, PSR.orm), lwd = 3)
abline(h = 0, lty = 2)


# residuals vs fitted value
plot_data <- data.frame(fitted = fitted, residuals = PSR.orm)

res.p3 <- ggplot(plot_data, aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "(c)", x = "Fitted Values", y = "Residuals")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))
res.p3

```

```{r}
# QQ-plot
qqplot(qunif(ppoints(length(PSR.orm)), -1, 1), PSR.orm, 
  main = " QQ-plots of probability-scale residuals (PSRs)", cex = 0.4, xlab = "Quantiles of Uniform Distribution",
  ylab = "Quantiles of PSRs") 
abline(a = 0, b = 1, lty = 2)
grid()

qq.p3 <- ggplot(plot_data, aes(sample = PSR.orm)) +
  geom_qq(distribution = function(p) qunif(p, min = -1, max = 1)) +
  geom_abline(slope = 1, intercept = 0,  linetype = "dashed") +
  labs(title = "(c)")+
  theme(plot.title = element_text(size = 6, hjust = 0.5))
qq.p3

```

### pooled result

```{r}
im <- aregImpute(~ daoohr30+hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, 
                 data = dt.im,
                 n.impute = 15)
dt.pooled.orm <- fit.mult.impute(daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp , xtrans = im, data = dt.im, fitter = orm)
orm_pool_result <- summary(dt.pooled.orm)
orm_pool_result
```

### mice + polr result

```{r}
dt.fit.ord <- with(dt.long.mids,
               polr(daoohr30.r~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, Hess=TRUE))

dt.pooled.ord <- pool(dt.fit.ord)
ord_pool_result <- summary(dt.pooled.ord)
```

### g-computation pooled result

```{r}
m <- 15
g_m <- numeric(m)  


# for each model, make prediction based on the each dataset and calculate ATE
for (i in 1:m) {
  dt.i <- impute.transcan(im, data=dt.im, imputation=i, list.out=TRUE, pr=FALSE, check=FALSE)

# convert the list to the database
  dt.i <- as.data.frame(dt.i, stringsAsFactors = FALSE)
  ddist <- datadist(dt.i)
  options(datadist = ddist)
  
  model.i <- orm(daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, x = TRUE, y = TRUE, data = dt.i)
  
  dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
  
  predicted.i0 <- predict(model.i,dt.i0,type = 'mean')
  
  dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
  predicted.i1 <- predict(model.i,dt.i1,type = 'mean')
  
  
  g_m[i] <- mean(predicted.i1 - predicted.i0) 
}

mean(g_m)
```

## Probabilistic Index Model

reference: 1. https://cran.r-project.org/web/packages/pim/vignettes/pim.pdf 2. De Schryver M, De Neve J. A tutorial on probabilistic index models: Regression models for the effect size P(Y1 \< Y2). Psychol Methods. 2019 Aug;24(4):403-418. doi: 10.1037/met0000194. Epub 2018 Sep 27. PMID: 30265047.(Wilcoxon–Mann–Whitney (WMW) test, t test, PIM)

```{r}
model.pim <- pim(daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, data = dt2)
summary(model.pim)
```

### Estimation and Interpretation

```{r}
exp(model.pim@coef['treatgroupSSU'])/(1 + exp(model.pim@coef['treatgroupSSU']))
```

The expit coefficient for treatgroupSSU is 0.60, which means given other covariates, the probability that subjects in the SSU treatment group have a higher daoohr30 compared to subjects in the Hospitalization group is estimated by 0.60.

### pooled result manually based on Rubin’s Rules

```{r}
pim_pool <- function(imp, formula, variable = "treatgroupSSU") {
  m <- imp$m
  q_m <- numeric(m)
  u_m <- numeric(m)
  
  for (i in 1:m) {
    dt.i <- complete(imp, i)
    formula_obj <- as.formula(formula)
    
    model <- pim(formula_obj, data = dt.i)

    coef <- model@coef[variable]
    se <- summary(model)@se[variable]
    
    q_m[i] <- coef
    u_m[i] <- se^2
    
  }
  
  # pooling
  Q_bar <- mean(q_m)
  U_bar <- mean(u_m)
  B <- var(q_m)
  T_var <- U_bar + (1 + 1/m) * B
  se_pooled <- sqrt(T_var)
  t_value <- Q_bar / se_pooled
  df <- (m - 1) * (1 + U_bar / ((1 + 1/m) * B))^2
  p_value <- 2 * pt(-abs(t_value), df)
  

  result <- list(
    estimate = Q_bar,
    std.error = se_pooled,
    t.value = t_value,
    df = df,
    p.value = p_value
  )
  
  return(result)
}

```

```{r}
pim_pool_result <- pim_pool(imp  = dt.imp, formula = 'daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp')
pim_pool_result
```

### compare with WMW test

```{r}
# same p-val as Wilcoxon–Mann–Whitney test
summary(pim(daoohr30~ treatgroup, data = dt2))
```

# G-computation CI

```{r}
#| echo: false
#| results: hide
library(Hmisc)
library(ggplot2)
library(rms)
library(mice)
library(boot)
library(MASS)
library(gt)
```

```{r}
# set the seed
set.seed(123)
```

## Functions

```{r}
bootstrap_data <- function(data) {
  
  n <- nrow(data)
  boot_data <- data[sample(1:n, size = n, replace = TRUE), ]
  
  return(boot_data)
  
}

glm_g_comp_one_sample_linear <- function(sample_imp_data, pool_model) {
  
  m <- m
  g_m <- numeric(m)  # store g-computation result
  
  # for each model, make prediction based on the each dataset and calculate ATE
  for (i in 1:m) {
    dt.i <- complete(sample_imp_data,i)
    dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
    predicted.i0 <- predict(pool_model$analyses[[i]],dt.i0)
    
    predicted.i0[predicted.i0 <0] <- 0
    predicted.i0[predicted.i0 >30] <- 30
    
    dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
    predicted.i1 <- predict(pool_model$analyses[[i]],dt.i1)
    
    predicted.i1[predicted.i1 <0] <- 0
    predicted.i1[predicted.i1 >30] <- 30
    
    # check values
    if (any(predicted.i0 < 0)) {
      print('negative values')
    } else if(any(predicted.i0 > 30)) {
      print('>30 values')
    } 
    
    if (any(predicted.i1 < 0)) {
      print('negative values1')
    } else if(any(predicted.i1 > 30)) {
      print('>30 values1')
    }
    
    g_m[i] <- mean(predicted.i1 - predicted.i0) 
  }
  
  return(g_m)
  
}

glm_g_comp_one_sample_gamma <- function(sample_imp_data, pool_model) {
  
  m <- m
  g_m <- numeric(m)  # store g-computation result
  
  # for each model, make prediction based on the each dataset and calculate ATE
  for (i in 1:m) {
    dt.i <- complete(sample_imp_data,i)
    dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
    predicted.i0 <- predict(pool_model$analyses[[i]],dt.i0)
    
    predicted.i0[predicted.i0 > log(30)] <- log(30)
    
    dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
    
    predicted.i1 <- predict(pool_model$analyses[[i]],dt.i1)
    
    predicted.i1[predicted.i1 > log(30)] <- log(30)
    
    # check values
    if (any(predicted.i0 > log(30))) {
      print('>30 values')
    }
    
    if (any(predicted.i1 > log(30))) {
      print('>30 values1')
    }
    
    
    g_m[i] <- mean(30 - exp(predicted.i1) - (30 - exp(predicted.i0))) 
  }
  
  return(g_m)
  
}

glm_g_comp_one_sample_ordinal <- function(sample_imp_data, pool_model) {
  
  m <- m
  g_m <- numeric(m)  # store g-computation result
  
  # for each model, make prediction based on the each dataset and calculate ATE
  for (i in 1:m) {
    dt.i <- impute.transcan(pool_model, data=sample_imp_data, imputation=i, list.out=TRUE, pr=FALSE, check=FALSE)


    dt.i <- as.data.frame(dt.i, stringsAsFactors = FALSE)
    ddist <- datadist(dt.i)
    options(datadist = ddist)
    
    model.i <- orm(daoohr30~ hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp, x = TRUE, y = TRUE, data = dt.i)
    
    dt.i0 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('Hospitalization', nrow(dt.i)))
    
    predicted.i0 <- predict(model.i,dt.i0,type = 'mean')
    
    dt.i1 <- cbind(dt.i[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval')], treatgroup = rep('SSU', nrow(dt.i)))
    predicted.i1 <- predict(model.i,dt.i1,type = 'mean')
    
    
    g_m[i] <- mean(predicted.i1 - predicted.i0)
    
  }
  
  return(g_m)
  
}


# store the result of all bootstrap result (n times of bootstrap) at original scale
g_comp_glm_result_linear <- function(n, m, family, data){
  
  result <- matrix(NA, nrow = n, ncol = m) #store the original scale result
  
  for (j in 1:n) {
    # bootstrap
    sample_data <- bootstrap_data(data)
    
    # multiple imputation
    suppressMessages(suppressWarnings(
      invisible(capture.output({
        sample_imp_data <- mice(sample_data, m = m, quiet = TRUE)
        }))
      ))
    
    sample_imp_long <- complete(sample_imp_data, action="long", include = TRUE)
    sample_imp_long$daoohr30 <- sample_data$daoohr30
    sample.mids <- as.mids(sample_imp_long)
    
    # fit the model
    pool_model <- with(sample.mids,
                       glm(formula = daoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup,
                           family = family))
    result[j, ] <- glm_g_comp_one_sample_linear(sample_imp_data,pool_model)
    
  }
  
  result_df <- as.data.frame(result)
  effect_of_interest <- rowMeans(result, na.rm = TRUE)
  
  return(effect_of_interest)
  
}


g_comp_glm_result_gamma <- function(n, m, family, data){
  
  result <- matrix(NA, nrow = n, ncol = m) #store the original scale result
  
  for (j in 1:n) {
    # bootstrap
    sample_data <- bootstrap_data(data)
    
    # multiple imputation
    suppressMessages(suppressWarnings(
      invisible(capture.output({
        sample_imp_data <- mice(sample_data, m = m, quiet = TRUE)
        }))
      ))
    
    sample_imp_long <- complete(sample_imp_data, action="long", include = TRUE)
    sample_imp_long$ndaoohr30 <- 30-sample_data$daoohr30
    sample.mids <- as.mids(sample_imp_long)
    
    # fit the model
    pool_model <- with(sample.mids,
                       glm(formula = ndaoohr30 ~ hosp6mo+sbp+bun+troponinilow2+bnpval+treatgroup,
                           family = family))
    result[j, ] <- glm_g_comp_one_sample_gamma(sample_imp_data,pool_model)
    
  }
  
  result_df <- as.data.frame(result)
  effect_of_interest <- rowMeans(result, na.rm = TRUE)
  
  return(effect_of_interest)
  
}

g_comp_glm_result_ordinal <- function(n, m, data){
  
  result <- matrix(NA, nrow = n, ncol = m) #store the original scale result
  
  for (j in 1:n) {
    # bootstrap
    sample_data <- bootstrap_data(data)
    
    # multiple imputation
    suppressMessages(suppressWarnings(
      invisible(capture.output({
        im <- aregImpute(~ daoohr30+hosp6mo+bun+troponinilow2+bnpval+treatgroup+sbp,
                     n.impute=m,
                     data = sample_data)
        }))
      ))
    
    
    result[j, ] <- glm_g_comp_one_sample_ordinal(sample_imp_data = sample_data,pool_model=im)
    
  }
  
  result_df <- as.data.frame(result)
  effect_of_interest <- rowMeans(result, na.rm = TRUE)
  
  return(effect_of_interest)
  
}

```

## Arguments

```{r}
m = 15
n = 500
```

## Data Preparation

```{r}
# new analysis data
dt2 <- read.csv('analytic_data.csv')

# transform data
dt2$daooh30 <- as.numeric(dt2$daooh30)
dt2$daoohr30 <- as.numeric(dt2$daoohr30)
dt2$daooh90 <- as.numeric(dt2$daooh90)
dt2$daoohr90 <- as.numeric(dt2$daoohr90)

dt2$hosp6mo <- as.factor(dt2$hosp6mo)
dt2$troponinilow2 <- as.factor(dt2$troponinilow2)
```

```{r}
dt2 <- dt2[!(is.na(dt2$daoohr30) & is.na(dt2$daoohr90)),]
dt2 <- dt2[,c('hosp6mo','sbp','bun', 'troponinilow2','bnpval','treatgroup','daoohr30')]
```

## CI calculation for linear regression

```{r}
# n, m, family, formula, data

linear_result <- g_comp_glm_result_linear(n = n,
                                          m = m,
                                          family = gaussian(link = "identity"), 
                                          data = dt2)

g_comp_linear <- quantile(linear_result,c(0.025,0.5,0.975))
g_comp_linear
```

## CI calculation for gamma regression

```{r}
# n, m, family, formula, data
gamma_result <- g_comp_glm_result_gamma(n = n,
                                        m = m,
                                        family = Gamma(link = "log"), 
                                        data = dt2)

g_comp_gamma <- quantile(gamma_result,c(0.025,0.5,0.975))
g_comp_gamma
```

## CI calculation for ordinal regression

```{r}
# n, m, family, formula, data

ordinal_result <- g_comp_glm_result_ordinal(n = n,
                                  m = m, 
                                  data = dt2)

g_comp_ord <- quantile(ordinal_result,c(0.025,0.5,0.975))
g_comp_ord
```

# Summary table for result

## Result Table 1: estimates, CI, p-val, estimand

```{r}
# collect the estimates

trt30r.estimate <- c(NA,
                     t.estimate,
                     linear_pool_result[7,'estimate'], # result of treatgroupSSU
                     NA,
                     NA,
                     exp(gamma_pool_result[7,'estimate']),
                     exp(orm_pool_result['treatgroup - SSU:Hospitalization','Effect']),
                     NA,
                     NA,
                     exp(wmw.estimate)/(1 + exp(wmw.estimate)),
                     exp(pim_pool_result$estimate)/(1 + exp(pim_pool_result$estimate))
                     )

trt30r.estimand <- c(NA,
                     'Mean',
                     'Mean',
                     NA,
                     NA,
                     'Geometreic Mean',
                     'Odds Ratio',
                     NA,
                     NA,
                     'Probabilistic Index',
                     'Probabilistic Index'
                     )

trt30r.lw <- c(NA,
               t.ci.lw,
               linear_pool_result[7,'estimate']-1.96*linear_pool_result[7,'std.error'],
               NA,
               NA,
               exp(gamma_pool_result[7,'estimate']-1.96*gamma_pool_result[7,'std.error']),
               exp(orm_pool_result['treatgroup - SSU:Hospitalization','Lower 0.95']),
               NA,
               NA,
               wmw.ci.lw,
               exp(pim_pool_result$estimate-1.96*pim_pool_result$std.error)/(1+exp(pim_pool_result$estimate-1.96*pim_pool_result$std.error))
               )

trt30r.up <- c(NA,
               t.ci.up,
               linear_pool_result[7,'estimate']+1.96*linear_pool_result[7,'std.error'],
               NA,
               NA,
               exp(gamma_pool_result[7,'estimate']+1.96*gamma_pool_result[7,'std.error']),
               exp(orm_pool_result['treatgroup - SSU:Hospitalization','Upper 0.95']),
               NA,
               NA,
               wmw.ci.up,
               exp(pim_pool_result$estimate+1.96*pim_pool_result$std.error)/(1+exp(pim_pool_result$estimate+1.96*pim_pool_result$std.error))
               )

# p_val for ordinal regression
L <- orm_pool_result['treatgroup - SSU:Hospitalization','Lower 0.95']
U <- orm_pool_result['treatgroup - SSU:Hospitalization','Upper 0.95']
SE <- orm_pool_result['treatgroup - SSU:Hospitalization','S.E.']
Est <- orm_pool_result['treatgroup - SSU:Hospitalization','Effect']
z <- Est/SE
ord_p_val <- 2 * (1 - pnorm(abs(z)))


trt30r.pval <- c(NA,
                 t.pval,
                 linear_pool_result[7,'p.value'],
                 NA,
                 NA,
                 gamma_pool_result[7,'p.value'],
                 ord_p_val, 
                 NA,
                 NA,
                 wmw.p.val,
                 pim_pool_result$p.value
                 )


trt30r.label <- c("Additive Model","T-test","Linear",'',
                  "Multiplicative Model","Gamma","Ordinal",'',
                  "Probabilistic Model","WMW-test","PIM")

# prepare the dataframe for plot
(dt.trt30r <- data.frame(subgroup = trt30r.label,
                         estimate = round(trt30r.estimate,2),
                         estimand = trt30r.estimand,
                         ci.low = round(trt30r.lw,2),
                         ci.up = round(trt30r.up,2),
                         pval = round(trt30r.pval,3)))
```

```{r}
# add the blank column to display CI

dt.trt30r$" " <- paste(rep(" ", 20), collapse = " ")

dt.trt30r$`95% CI` <- ifelse((is.na(dt.trt30r$ci.low)), "",
                             sprintf("(%.2f to %.2f)",
                                     dt.trt30r$ci.low, dt.trt30r$ci.up))

dt.trt30r$estimate <- ifelse(is.na(dt.trt30r$estimate),"",dt.trt30r$estimate)
dt.trt30r$ci.low <- ifelse(is.na(dt.trt30r$ci.low),"",dt.trt30r$ci.low)
dt.trt30r$ci.up <- ifelse(is.na(dt.trt30r$ci.up),"",dt.trt30r$ci.up)
dt.trt30r$pval <- ifelse(is.na(dt.trt30r$pval),"",dt.trt30r$pval)
dt.trt30r$estimand <- ifelse(is.na(dt.trt30r$estimand),"",dt.trt30r$estimand)

dt.trt30r$subgroup <- ifelse(is.na(dt.trt30r$estimate), dt.trt30r$subgroup,paste0("   ", dt.trt30r$subgroup))
dt.trt30r[c('subgroup','estimate','estimand','pval','95% CI')]
```

```{r}
dt.trt30r[c('subgroup','estimate','estimand','pval','95% CI')] %>%
  gt() %>%
  tab_header(title = "Table 1") %>%
  fmt_number(columns = c("pval"), decimals = 2) %>%
  cols_label(
    subgroup = "Subgroup",
    estimate = "Estimate",
    estimand = "Estimand",
    pval = "P-value",
    '95% CI' = "95%CI"
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(
      rows = subgroup %in% c("Additive Model", "Multiplicative Model"),
      columns = everything()
    )
  )
```

## Result Table 2: g-computation result: linear, gamma, ord

```{r}
# collect the estimates

g.comp.estimate <- c(NA,
                     g_comp_linear[2],
                     NA,
                     NA,
                     g_comp_gamma[2],
                     g_comp_ord[2]
                     )


g.comp.lw <- c(NA,
               g_comp_linear[1],
               NA,
               NA,
               g_comp_gamma[1],
               g_comp_ord[1]
               )

g.comp.up <- c(NA,
               g_comp_linear[3],
               NA,
               NA,
               g_comp_gamma[3],
               g_comp_ord[3]
               )



g.comp.label <- c("Additive Model","Linear",'',
                  "Multiplicative Model","Gamma","Ordinal")

# prepare the dataframe for plot
(dt.g.comp <- data.frame(subgroup = g.comp.label,
                         estimate = round(g.comp.estimate,2),
                         ci.low = round(g.comp.lw,2),
                         ci.up = round(g.comp.up,2)))
```

```{r}
dt.g.comp$" " <- paste(rep(" ", 20), collapse = " ")


dt.g.comp$`95% CI` <- ifelse((is.na(dt.g.comp$ci.low)), "",
                             sprintf("(%.2f to %.2f)",
                                     dt.g.comp$ci.low, dt.g.comp$ci.up))

dt.g.comp$estimate <- ifelse(is.na(dt.g.comp$estimate),"",dt.g.comp$estimate)

dt.g.comp$subgroup <- ifelse(is.na(dt.g.comp$estimate), dt.g.comp$subgroup,paste0("   ", dt.g.comp$subgroup))
dt.g.comp[c('subgroup','estimate','95% CI')]
```

```{r}
dt.g.comp[c('subgroup','estimate','95% CI')] %>%
  gt() %>%
  tab_header(title = "Table 2") %>%
  cols_label(
    subgroup = "Subgroup",
    estimate = "Estimate",
    '95% CI' = "95%CI"
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(
      rows = subgroup %in% c("Additive Model", "Multiplicative Model"),
      columns = everything()
    )
  )
```

## Diagnostic Plots

```{r}
res.p1.n <- res.p1 + labs(y = 'Residuals',x = 'DAOOH') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
res.p2.n <- res.p2 + labs(y = 'Residuals',x = 'DAOOH(Rank transformed)') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
res.p3.n <- res.p3 + labs(y = 'PResiduals',x = 'DAOOH') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))

res.plot <- ggarrange(res.p1.n, res.p2.n,res.p3.n,
                       ncol = 3, nrow = 1,
                       align = "hv",
                       widths = c(1, 1),
                       common.legend = FALSE)
res.plot <- annotate_figure(res.plot)
res.plot
ggsave("res.plot.png", plot = res.plot, width = 10, height = 3, dpi = 300)
```

```{r}
qq.p1.n <- qq.p1 + labs(y = NULL, x = 'Quantiles of Normal Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
qq.p2.n <- qq.p2 + labs(y = NULL, x = 'Quantiles of Uniform Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
qq.p3.n <- qq.p3 + labs(y = NULL, x = 'Quantiles of Uniform Distribution') + theme_bw()+
  theme(plot.title = element_text(size = 8, hjust = 0.5))
qq.plot <- ggarrange(qq.p1.n, qq.p2.n,qq.p3.n,
                       ncol = 3, nrow = 1,
                       align = "hv",
                       widths = c(1, 1),
                       common.legend = FALSE)
qq.plot <- annotate_figure(qq.plot,
                         left   = text_grob("Sample Quantiles", rot = 90, size = 10),
                         bottom = text_grob("Theoretical Quantiles", size = 10))
qq.plot
ggsave("qq.plot.png", plot = qq.plot, width = 10, height = 3, dpi = 600)
```
